Open Data, Open Source and Open Standards in chemistry: \\ The Blue Obelisk five years on

### 2011-08-16 paper.tex
Proper spelling of väg.
Response to reviewers. Note: a little more work required for reviewer 2, but I think reviewer 1 is finished

\address{%
    \iid(1)Analytical and Biological Chemistry Research Facility, Cavanagh Pharmacy Building, University College Cork, College Road, Cork, Co. Cork, Ireland \\
    \iid(2)NIH Center for Translational Therapeutics, 9800 Medical Center Drive, Rockville, MD 20878, USA \\
    \iid(3)Division of Molecular Toxicology, Institute of Environmental Medicine, %
        Nobels vaeg 13, Karolinska Institutet, 171 77 Stockholm, Sweden \\
    \iid(4)Unilever Centre for Molecular Sciences Informatics, Department of Chemistry, University of Cambridge, Lensfield Road, CB2 1EW, UK \\
    \iid(5)Department of Pharmaceutical Biosciences, Uppsala University, Box 591, 751 24 Uppsala, Sweden \\
    \iid(6)Department of Chemistry, Drexel University, 32nd and Chestnut streets, Philadelphia, PA 19104, USA \\
    \iid(7)Chemical Biology Laboratory, Basic Research Program, SAIC-Frederick,
Inc., NCI-Frederick, Frederick, MD 21702, USA \\
    \iid(8)St. Olaf College, 1520 St. Olaf Ave., Northfield, MN 55057, USA \\
    \iid(9)Kitware, Inc., 28 Corporate Drive, Clifton Park, NY 12065, USA \\
    \iid(10)Department of Chemistry, University of Pittsburgh, 219 Parkman Avenue, Pittsburgh, PA 15260, USA \\
    \iid(11)eMolecules Inc., 380 Stevens Ave., Solana Beach, California 92075, USA \\
    \iid(12)Ideaconsult Ltd., 4.A.Kanchev str., Sofia 1000, Bulgaria \\
    \iid(13)Department of Engineering, Computer Science, Physics, and Mathematics, Oral Roberts University, 7777 S. Lewis Ave. Tulsa, OK 74171, USA \\
    \iid(14)Leiden Institute of Chemistry, Leiden University, Einsteinweg 55, 2333 CC Leiden, The Netherlands \\
    \iid(15)Department of Chemistry, State University of New York at Buffalo, Buffalo, NY 14260-3000, USA \\
    \iid(16)Universit\'{e} de Strasbourg, IPHC, CNRS, UMR7178, 23 rue du Loess 67037, Strasbourg, France \\
    \iid(17)GGA Software Services LLC, 41 Nab. Chernoi rechki 194342, Saint Petersburg, Russia \\
    \iid(18)Cheminformatics and Metabolism Team, European Bioinformatics Institute (EBI), Wellcome Trust Genome Campus, Hinxton, Cambridge, UK \\
    \iid(19)Department of Chemistry, University of Washington, Seattle, WA 98195, USA \\
    \iid(20)iChemLabs, 200 Centennial Ave., Suite 200, Piscataway, NJ 08854, USA 
}%

@@

        \paragraph*{Results:} This
		contribution looks back on the past 5 years and surveys progress and
		remaining challenges in the areas of Open Data, Open Standards, and
    Open Source in chemistry.

@@

The Blue Obelisk was first described in the CDK News \cite{CDKNewsBO} and
later as a formal paper by Guha et al.\cite{Guha2006} in
2006. Its home on the web is at http://blueobelisk.org.
This contribution looks back on the past 5 years and surveys
progress and remaining challenges in the areas of Open Data, Open
Source, and Open Standards in chemistry.

@@

Besides the traditional and proven relational database approach with
added chemical features (`cartridges'), there is growing
interest in tools and approaches based on the web philosophy and
practice. Several groups are experimenting with 
the Resource Description Framework (RDF) language
on the assumption that generic high-performance solutions will appear.
RDF allows everything to be described by
URIs (data, molecules, dictionaries, relations). The Chempound
system~\cite{Chempound},
as deployed in
Quixote and elsewhere, is an RDF-based approach to chemical structures
and compounds and their properties. For small
to medium-sized collections (such as an individual's calculations or
literature retrieval), there are many RDF tools
(e.g. SIMILE, Apache Jena) which can operate in machine memory and provide
the flexibility that RDF offers. For larger
systems, it is unclear whether complete RDF solutions (e.g. Virtuoso)
will be satisfactory or whether a hybrid system
based on name-value pairs (e.g. CouchDB, MongoDB) will be sufficient.

>>>>

\address{%
    \iid(1)Analytical and Biological Chemistry Research Facility, Cavanagh Pharmacy Building, University College Cork, College Road, Cork, Co. Cork, Ireland \\
    \iid(2)NIH Center for Translational Therapeutics, 9800 Medical Center Drive, Rockville, MD 20878, USA \\
    \iid(3)Division of Molecular Toxicology, Institute of Environmental Medicine, %
        Nobels v\"ag 13, Karolinska Institutet, 171 77 Stockholm, Sweden \\
    \iid(4)Unilever Centre for Molecular Sciences Informatics, Department of Chemistry, University of Cambridge, Lensfield Road, CB2 1EW, UK \\
    \iid(5)Department of Pharmaceutical Biosciences, Uppsala University, Box 591, 751 24 Uppsala, Sweden \\
    \iid(6)Department of Chemistry, Drexel University, 32nd and Chestnut streets, Philadelphia, PA 19104, USA \\
    \iid(7)Chemical Biology Laboratory, Basic Research Program, SAIC-Frederick,
Inc., NCI-Frederick, Frederick, MD 21702, USA \\
    \iid(8)St. Olaf College, 1520 St. Olaf Ave., Northfield, MN 55057, USA \\
    \iid(9)Kitware, Inc., 28 Corporate Drive, Clifton Park, NY 12065, USA \\
    \iid(10)Department of Chemistry, University of Pittsburgh, 219 Parkman Avenue, Pittsburgh, PA 15260, USA \\
    \iid(11)eMolecules Inc., 380 Stevens Ave., Solana Beach, California 92075, USA \\
    \iid(12)Ideaconsult Ltd., 4.A.Kanchev str., Sofia 1000, Bulgaria \\
    \iid(13)Department of Engineering, Computer Science, Physics, and Mathematics, Oral Roberts University, 7777 S. Lewis Ave. Tulsa, OK 74171, USA \\
    \iid(14)Leiden Institute of Chemistry, Leiden University, Einsteinweg 55, 2333 CC Leiden, The Netherlands \\
    \iid(15)Department of Chemistry, State University of New York at Buffalo, Buffalo, NY 14260-3000, USA \\
    \iid(16)Universit\'{e} de Strasbourg, IPHC, CNRS, UMR7178, 23 rue du Loess 67037, Strasbourg, France \\
    \iid(17)GGA Software Services LLC, 41 Nab. Chernoi rechki 194342, Saint Petersburg, Russia \\
    \iid(18)Cheminformatics and Metabolism Team, European Bioinformatics Institute (EBI), Wellcome Trust Genome Campus, Hinxton, Cambridge, UK \\
    \iid(19)Department of Chemistry, University of Washington, Seattle, WA 98195, USA \\
    \iid(20)iChemLabs, 200 Centennial Ave., Suite 200, Piscataway, NJ 08854, USA 
}%

@@

        \paragraph*{Results:} This
		contribution looks back on the work carried out by the
Blue Obelisk in the past 5 years and surveys progress and
		remaining challenges in the areas of Open Data, Open Standards, and
    Open Source in chemistry.

@@

The Blue Obelisk was first described in the CDK News \cite{CDKNewsBO} and
later as a formal paper by Guha et al.\cite{Guha2006} in
2006. Its home on the web is at http://blueobelisk.org.
This contribution looks back on the work carried out by the Blue
Obelisk over the past 5 years in the areas of Open Data, Open
Source, and Open Standards in chemistry.

@@

Besides the traditional and proven relational database approach with
added chemical features (`cartridges'), there is growing
interest in tools and approaches based on the web philosophy and
practice. Several groups~\cite{willighagen_resource_2011,chen_chem2bio2rdf_2010} are experimenting with 
the Resource Description Framework (RDF) language
on the assumption that generic high-performance solutions will appear.
RDF allows everything to be described by
URIs (data, molecules, dictionaries, relations). The Chempound
system~\cite{Chempound},
as deployed in
Quixote and elsewhere, is an RDF-based approach to chemical structures
and compounds and their properties. For small
to medium-sized collections (such as an individual's calculations or
literature retrieval), there are many RDF tools
(e.g. SIMILE, Apache Jena) which can operate in machine memory and provide
the flexibility that RDF offers. For larger
systems, it is unclear whether complete RDF solutions (e.g. Virtuoso)
will be satisfactory or whether a hybrid system
based on name-value pairs (e.g. CouchDB, MongoDB) will be sufficient.
